---
title: "02_PP_all_reps_subset_and_markers"
author: "Ananya Gupta"
date: "2025-11-14"
output: html_document
---

```{r env_loading, results='FALSE',}
# Load packages, data and functions
library(Seurat)
library(tidyverse)
library(SeuratWrappers)
library(magicBatch)
library(BPCells)
library(batchelor)
library(patchwork)
library(Azimuth)
library(plotly)
library(gprofiler2)
library(hdf5r)
library(kableExtra)
library(ggplot2)
library(gridExtra)
library(DT)
library(clustree)
library(Matrix)
library(scales)
library(data.table)
library(dplyr)
library(tidyverse)
library(glue)
library(ggrepel)
library(gprofiler2)
library(multtest)
library(metap)

#for parallelization
options(future.globals.maxSize = 3000 * 1024^2) 
#options(future.globals.maxSize = 1e9) was abt 1GB


# Set the random number seed
set.seed(1234)
```


```{r dir setup, include=FALSE}
base_dir <- "PP_outputs"

# Define output directories
rds_dir  <- file.path(base_dir, "PP_all_rep_rds")
fig_dir  <- file.path(base_dir, "PP_all_rep_figs")

# Create folders if they don't exist
dir.create(base_dir, showWarnings = FALSE)
dir.create(rds_dir, showWarnings = FALSE)
dir.create(fig_dir, showWarnings = FALSE)
```
 saving a fig ggsave(
  filename = file.path(fig_dir, "cluster_umap.png"),
  plot = p,
  width = 6,
  height = 5,
  dpi = 300
)


saveRDS(
  object = seu,
  file = file.path(rds_dir, "seu_pp_all_reps.rds")
)


```{r}
library(Cairo)
save_pub <- function(plot,
                     filename,
                     width = 7,
                     height = 7,
                     dpi = 600) {

  

  # SAVE PDF
  ggsave(
    filename = file.path(fig_dir, paste0(filename, ".pdf")),
    plot = plot,
    device = CairoPDF,
    width = width,
    height = height,
    units = "in"
  )

  # SAVE PNG 
  CairoPNG(
    filename = file.path(fig_dir, paste0(filename, ".png")),
    width  = width * dpi,   # convert inches → pixels
    height = height * dpi,
    res    = dpi
  )

  print(plot) 
  dev.off()

  return(plot)
}

```

```{r}
# Load it back
#seu <- readRDS("seu_int.rds")
seu <- readRDS(file = file.path(rds_dir,"seu_pp_int_cc_calc.rds"))
```

you might have to do cc diff if you want it cause id idnt sve that



```{r}
p <- ggplot(seu@meta.data) +
    geom_bar(aes(x=orig.ident, fill=condition),
             stat="count", color="black") +
    theme_classic() +
    theme(axis.text.x = element_text(angle=45, vjust=1, hjust=1)) +
    labs(x="Sample name", y="Number of cells")
save_pub(plot=p, filename="cellno")
```

```{r}
p <- DimPlot(seu, group.by="seurat_clusters") + NoLegend()
LabelClusters(p, id = "seurat_clusters",  fontface = "bold", size = 5, bg.colour = "white", bg.r = .2, force = 0)
```

Okay so these many clusters aree too much for me rn. We need 7-8 sow e will try again.Ill make it into a new object since I get bothered when its the same one.


```{r}
seu_obj <- FindClusters(seu, resolution = seq(0.1, 0.4, by=0.1))
```
```{r Find best cluster resolution,fig.height=8,fig.width=10}

clustree::clustree(seu_obj@meta.data[,grep("RNA_snn_res.", colnames(seu_obj@meta.data))],prefix = "RNA_snn_res.", layout="sugiyama")+ggtitle("Finding Best Resolution")+
  theme(
    plot.title = element_text(
      hjust = 0.5,       # centers the title
      margin = margin(b = 12)  # adds space (bottom margin)
    )
  )
```
Let's try with 0.2 and see if I am into that.How? Well we should first have a marker list, derived from https://www.nature.com/articles/s41467-020-17291-5#Sec2
We will remove contaminats too which are also imp so I will include that as well for now.

Clusters extending to the termini of the arterial, CRP, and HEV branches were further separated to distinguish cells that were most distinct (distant along trajectories) from the bulk of EC (darker shading, Fig. 1e): these cells are enriched for genes associated with mature arterial (e.g., Gkn3, Bmx) or HEV (Chst4, Glycam1) differentiation, or in the case of CRP, for genes associated with endothelial specification during early development or with stem cells (e.g., Ets2, Cxcr4, Nes) 

```{r}
#Contaminants
lec <- c("Prox1","Lyve1","Pdpn")
lymphocytes <- c("Ptprc","Cd52")
fib_ret <- c("Pdpn","Ccl19","Pdgfra")

# cell markers

hev_markers <- c("St6gal1", "Glycam1", "Chst4", "Ccl21a", "Madcam1")
art_markers <- c("Bmx", "Depp1", "Gja4", "Gja5", "Gkn3", "Sox17")
CRP_markers <- c("Angpt2", "Apln", "Esm1", "Mcam", "Nid2", "Pdgfb", "Pgf", "Vim")
PCV_markers <- c("Ackr1", "Icam1", "Nr2f2", "Sele", "Selp", "Vcam1", "Vwf")
CapEC_marker <- c("Cdh13", "Emcn", "Gja1", "Gpihbp1", "Ly6a", "Ly6c1", "Podxl", "Ramp3")
CapEC2_marker <- c("Atf3", "Cxcl1", "Egr1", "Fos", "Fosb", "Jun", "Junb", "Jund", "Nfkbia", "Nr4a1", "Sgk1") #this overlaps w a lot of others and the capec1 too
CapEC1_marker <- c("Col4a1", "Col4a2", "Hlx", "Id1", "Id3", "Igfbp3")
art_vn_markers <- c("Edn1", "Eln", "Fbln5", "Foxn3", "Klf4", "Ltbp4", "Ptprb")



```


For 0.2 analysis
```{r}
seu_obj <- SetIdent(seu_obj, value = seu_obj$RNA_snn_res.0.2)
```

```{r,fig.height=15,fig.width=7}
Seurat::DimPlot(seu_obj, group.by = c("RNA_snn_res.0.1","RNA_snn_res.0.2", "RNA_snn_res.0.3"),ncol=1, label=T)
```


```{r}
p_0.2 <- DimPlot(seu_obj, group.by="RNA_snn_res.0.2") #+ NoLegend()
LabelClusters(p_0.2, id = "RNA_snn_res.0.2",  fontface = "bold", size = 5, bg.colour = "white", bg.r = .2, force = 0)
```

```{r,fig.width=7,fig.height=8}
FeaturePlot(seu_obj,lec,label=T) #3 and 4  #in 0.3 its 2 n 3 respectively
FeaturePlot(seu_obj,lymphocytes,label=T) #clust 9 #13 in 0.3
FeaturePlot(seu_obj,fib_ret,label=T) #3 and 4  
```

```{r,fig.height=8,fig.width=8}
FeaturePlot(seu_obj,art_markers,label=T)
FeaturePlot(seu_obj,CapEC1_marker,label=T)
```
```{r,fig.width=8,fig.height=9}
FeaturePlot(seu_obj,PCV_markers,label=T)
```
So from what I'm seeing 0.3 might be the best oen and we def cant take out clusters, only cells as im seeing clear HEC division in 0.3 unlike 0.2. and clusters have mult cells of inetrest so we will have to take them out.
```{r}
?IntegrateLayers
```










```{r}
DefaultAssay(seu) <- "RNA"

```

Lets try to find markers using the annotations way

```{r}
annotations <- read.csv("mouse_annotations.csv")
```

```{r}
head(annotations)
```
```{r}
head(seu@meta.data)
```


```{r}
get_conserved <- function(cluster){
  FindConservedMarkers(seu,
                       ident.1 = cluster,
                       grouping.var = "condition",
                       only.pos = TRUE) %>%
    rownames_to_column(var = "gene") %>%
    left_join(y = unique(annotations[, c("gene_name", "description")]),
               by = c("gene" = "gene_name")) %>%
    cbind(cluster_id = cluster, .)
  }
```

```{r}
#map_dfr(inputs_to_function, name_of_function)
levels(Idents(seu))
clusters_to_use <- levels(Idents(seu))
conserved_markers <- map_dfr(clusters_to_use, get_conserved)
```
```{r}
head(conserved_markers)
```


For now manually looking at the skipped clusters.Before that we evaluate

```{r}
top10_wt <- conserved_markers %>%
  group_by(cluster_id) %>%
  slice_max(order_by = WT_avg_log2FC, n = 10)

```


```{r}
top10 <- conserved_markers %>% 
  mutate(avg_fc = (WT_avg_log2FC + OE_avg_log2FC + KO_avg_log2FC) /3) %>% 
  group_by(cluster_id) %>% 
  top_n(n = 10, 
        wt = avg_fc)

# Visualize top 10 markers per cluster
View(top10)
```

```{r}
markers <- FindAllMarkers(object = seu, 
                          only.pos = TRUE,
                          logfc.threshold = 0.25)   
```
```{r}
get_cluster_markers <- function(df, cl, top = 10) {
  df %>%
    filter(cluster == cl) %>%
    arrange(desc(avg_log2FC)) %>%
    head(top)
}

get_cluster_markers(markers, 8, top = 20
                    )

```


```{r}
#print(top10[top10$cluster_id == 1, "gene"])
as.list(top10[top10$cluster_id == 4
              , "gene"])


```
```{r}
paste(top10[top10$cluster_id == 1, "gene"], collapse = ", ")

```
```{r}
#save cluster 
write.csv(top10, "output.csv", row.names = FALSE)


```

```{r}
# Determine differentiating markers for CD4+ T cell
cd4_tcells <- FindMarkers(seu,
                          ident.1 = 2,
                          ident.2 = c(0,4,6))                  

# Add gene symbols to the DE table
cd4_tcells <- cd4_tcells %>%
  rownames_to_column(var = "gene") %>%
  left_join(y = unique(annotations[, c("gene_name", "description")]),
             by = c("gene" = "gene_name"))

# Reorder columns and sort by padj      
cd4_tcells <- cd4_tcells[, c(1, 3:5,2,6:7)]

cd4_tcells <- cd4_tcells %>%
  dplyr::arrange(p_val_adj) 

# View data
View(cd4_tcells)


```



```{r}
genes <- c("Pgf", "Apln", "Hey1")
FeaturePlot(seu, features = genes, label = T)
```

```{r}
genes <- c("Chst4", "Ccl21a")
FeaturePlot(seu, features = genes, label = T)
```


```{r}
genes <- c("Selp", "Vcam1")
FeaturePlot(seu, features = genes, label = T)
```


```{r}
genes <- c("St6gal1")
FeaturePlot(seu, features = genes, label = T)

```
```{r}
genes <- c("Vwf", "Selp", "Ptn", "Ackr1")
FeaturePlot(seu, features = genes, label = T)
```

```{r}
#Art
hec_core <- c("Glycam1", "Madcam1", "Chst4", "Fut7", "Ackr4")
FeaturePlot(seu, features = hec_core)
#
```

```{r}
seu$celltype <- "Unknown"


```
```{r}
seu$celltype[seu$seurat_clusters == 0] <- "CapEC"
seu$celltype[seu$seurat_clusters == 3] <- "Art"
seu$celltype[seu$seurat_clusters == 5] <- "HEC"
seu$celltype[seu$seurat_clusters == 9] <- "Pericytes"
seu$celltype[seu$seurat_clusters == 11] <- "LEC"
seu$celltype[seu$seurat_clusters == 11] <- "LEC"

```

```{r}
c <- DimPlot(seu, group.by="seurat_clusters") + NoLegend()
LabelClusters(c, id = "seurat_clusters",  fontface = "bold", size = 5, bg.colour = "white", bg.r = .2, force = 0)
```
```{r}
head(seu@meta.data)
```
```{r}
seu$celltype_manual[seu$seurat_clusters == 0]  <- "CapEC"
seu$celltype_manual[seu$seurat_clusters == 1]  <- "Art"
seu$celltype_manual[seu$seurat_clusters == 4]  <- "PCV"
seu$celltype_manual[seu$seurat_clusters == 5]  <- "HEC"
seu$celltype_manual[seu$seurat_clusters == 8]  <- "Vn"
seu$celltype_manual[seu$seurat_clusters == 9]  <- "Pericytes"
seu$celltype_manual[seu$seurat_clusters == 11] <- "LEC"

```


```{r}
seu <- RenameIdents(object = seu, 
                               "0" = "CapEC",
                               "1" = "Art",
                               "2" = "X",
                               "3" = "X",
                               "4" = "PCV",
                               "5" = "HEC",
                               "6" = "X",
                               "7" = "X",
                               "8" = "Vn",
                               "9" = "Pericytes",
                               "10" = "X",
                               "11" = "LEC",
                               "12" = "X",
                               "13" = "X")



#Plot the UMAP
DimPlot(object = seu,
        reduction = "seurat_clusters",
        label = TRUE,
        label.size = 3,
        repel = TRUE)
```

```{r}
c <- DimPlot(seurat_integrated, group.by="seurat_clusters") + NoLegend()
LabelClusters(c, id = "celltype",  fontface = "bold", size = 5, bg.colour = "white", bg.r = .2, force = 0)
```

