---
title: "05_PP_CRP"
author: "Ananya Gupta"
date: "2025-11-26"
output: html_document
---


```{r env_loading, results='FALSE',}
# Load packages, data and functions
library(Seurat)
library(tidyverse)
library(SeuratWrappers)
library(magicBatch)
library(BPCells)
library(pheatmap)
library(batchelor)
library(patchwork)
library(plotly)
library(gprofiler2)
library(hdf5r)
library(kableExtra)
library(ggplot2)
library(gridExtra)
library(DT)
library(clustree)
library(Matrix)
library(scales)
library(data.table)
library(dplyr)
library(tidyverse)
library(glue)
library(msigdbr)
library(ggrepel)
library(gprofiler2)
library(multtest)
library(scater)
library(metap)
library(future)
library(parallelly)
library(scuttle)
library(SingleCellExperiment)
library(edgeR)
library(limma)
library(EnhancedVolcano)
library(fgsea)
library(KEGGREST)
library(pathview)
library(AnnotationDbi)

library(org.Mm.eg.db)
library(clusterProfiler)
plan(multicore, workers = 8)


#for parallelization
options(future.globals.maxSize = 8000 * 1024^2) 
#options(future.globals.maxSize = 1e9) was abt 1GB


# Set the random number seed
set.seed(1234)
```


```{r dir setup, include=FALSE}
base_dir <- "PP_outputs"

# Define output directories
rds_dir  <- file.path(base_dir, "PP_all_rep_rds")
fig_dir  <- file.path(base_dir, "PP_all_rep_figs")
celltype_dir <- file.path(base_dir, "PP_CapEC1")
dir.create(celltype_dir, showWarnings = FALSE, recursive = TRUE)


# Create folders if they don't exist
dir.create(base_dir, showWarnings = FALSE)
dir.create(rds_dir, showWarnings = FALSE)
dir.create(fig_dir, showWarnings = FALSE)
```

```{r}
library(Cairo)
save_pub <- function(plot,
                     filename,
                     width = 7,
                     height = 7,
                     dpi = 600) {

  

  # SAVE PDF
  ggsave(
    filename = file.path(celltype_dir, paste0(filename, ".pdf")),
    plot = plot,
    device = CairoPDF,
    width = width,
    height = height,
    units = "in"
  )

  # SAVE PNG 
  CairoPNG(
    filename = file.path(celltype_dir, paste0(filename, ".png")),
    width  = width * dpi,   # convert inches → pixels
    height = height * dpi,
    res    = dpi
  )

  print(plot) 
  dev.off()

  return(plot)
}


```


```{r}
seu <- readRDS(file.path(rds_dir, "seu_pp_7BEC_process_04res.rds"))
```


seu@meta.data$cluster_annot[seu@meta.data$seurat_clusters == 0] <- "CapEC2"
seu@meta.data$cluster_annot[seu@meta.data$seurat_clusters == 1] <- "Artery"
seu@meta.data$cluster_annot[seu@meta.data$seurat_clusters == 2] <- "Cap Ifn"
seu@meta.data$cluster_annot[seu@meta.data$seurat_clusters == 3] <- "CapEC1"
seu@meta.data$cluster_annot[seu@meta.data$seurat_clusters == 4] <- "PCV"
seu@meta.data$cluster_annot[seu@meta.data$seurat_clusters == 5] <- "HEV"
seu@meta.data$cluster_annot[seu@meta.data$seurat_clusters == 6] <- "Pre-Artery"
seu@meta.data$cluster_annot[seu@meta.data$seurat_clusters == 7] <- "Vn Sele-"
seu@meta.data$cluster_annot[seu@meta.data$seurat_clusters == 8] <- "Barrier EC"
seu@meta.data$cluster_annot[seu@meta.data$seurat_clusters == 9] <- "CRP"

Before doing DGE, you should aggregate (combine) the expression values of cells from the same sample and the same cell type — using either sum, mean, or a mixed-model random effect. This aggregation step is called pseudobulk, and it is necessary because cells inside the same sample are correlated (not independent).

Lets start with HEV


```{r,fig.width=14,fig.height=5}
an2 <- DimPlot(seu, split.by= "condition",group.by="cluster_annot") + ggtitle("Annotated Peyers Patches") +
  theme(plot.title = element_text(hjust = 0.5))
LabelClusters(an2, id = "cluster_annot",  fontface = "bold", size = 2.5, bg.colour = "white", bg.r = .2, force = 0)
an
```



```{r}
ets_family_genes <- c(
  "Ehf", "Elf1", "Elf2", "Elf3", "Elf4", "Elf5",
  "Elk1", "Elk3", "Elk4",
  "Erf", "Erg",
  "Ets1", "Ets2",
  "Etv1", "Etv2", "Etv4", "Etv5", "Etv6",
  "Fli1",
  "Gabpa", "Gabpb1", "Gabpb2",
  "Spdef",
  "Spi1", "Spib", "Spic"
)
# 
# all_paths <- keggList("pathway", "mmu")
# 
# ## 2. Turn each KEGG pathway into a gene-symbol set
# kegg_gs <- lapply(names(all_paths), function(pid) {
#   info <- keggGet(pid)[[1]]
#   genes <- info$GENE
#   if (is.null(genes)) return(NULL)
# 
#   # KEGG returns: c("ID1", "SYMBOL1; desc", "ID2", "SYMBOL2; desc", ...)
#   symbols <- genes[seq(2, length(genes), 2)]
#   symbols <- sapply(strsplit(symbols, ";"), `[`, 1)
#   trimws(symbols)
# })
# names(kegg_gs) <- names(all_paths)
# kegg_gs <- kegg_gs[!sapply(kegg_gs, is.null)]

#kegg_gs
```

kegg_gs is genes for each mouse pathway that we grab.
$mmu04670
[1] "Gene1" "Gene2" "Gene3" ...

$mmu04066
  [1] "Gm30155"  "Gm34005"  "Ldhal6b"  "Camk2d"   "Egln1"    "Egln2"    "Egln3"    "Gm29667"  "Angpt1"   "Angpt2"  
 [11] "Angpt4"   "Akt1"     "Akt2"     "Aldoa"    "Aldoc"    "Arnt"     "Bcl2"     "Camk2a"   "Camk2b"  
```{r}
#saveRDS(kegg_gs, file = "kegg_gs_mouse.rds")
kegg_gs <- readRDS("kegg_gs_mouse.rds")

```

```{r}
#saveRDS(msig_c3, file = "msig_c3.rds")
msig_c3 <- readRDS("msig_c3.rds")

```

Important to create cell type pseudobulk


```{r}
pseudobulk_celltype <- function(
  seu, 
  celltype, 
  n_pseudo = 2,
  assay_name = "RNA",
  oe_samples = c("COUPOE-PP1", "COUPOE-PP2"),
  seed = 1234
) {
   
  #----------------------
  # 1. Subset Seurat to that cell type
  #----------------------
  seu_sub <- subset(seu, subset = orig.ident != "COUPOE-PP-POS1")
  seu_sub <- subset(seu_sub, subset = cluster_annot == celltype)
  message("Cells in this cell type: ", ncol(seu_sub))
  if (ncol(seu_sub) < 50) warning("Very few cells — PCA & DGE may be unstable.")
  
  #----------------------
  # 2. Convert to SCE
  #----------------------
  sce <- as.SingleCellExperiment(seu_sub, assay = assay_name)
  
  
  meta <- as.data.frame(colData(sce))
  meta$cell <- rownames(meta)
  
  set.seed(seed)
  
  
meta_out <- meta %>%
  group_by(orig.ident) %>%
  do({
    df <- .
    
    if (df$orig.ident[1] %in% oe_samples) {
      # Assign pseudo_rep = position in oe_samples
      rep_index <- match(df$orig.ident[1], oe_samples)
      df$pseudo_rep <- rep_index
    } else {
      # Regular random pseudoreps for WT, KO, CRE_CTRL
      df$.rand <- sample(seq_len(nrow(df)))
      df$pseudo_rep <- rep(1:n_pseudo, length.out = nrow(df))[order(df$.rand)]
    }
    
    df
  }) %>%
  ungroup()

  colData(sce)$pseudo_rep <- meta_out$pseudo_rep[match(colnames(sce), meta_out$cell)]

  
  #colData(sce)$pseudo_rep <- meta$pseudo_rep[match(colnames(sce), meta$cell)]
  
  #----------------------
  # 4. Make pseudobulk IDs
  #----------------------
  # colData(sce)$pb_id <- with(
  #   colData(sce),
  #   paste(orig.ident, paste0("rep", pseudo_rep), sep = "_")
  # )
  colData(sce)$pb_id <- ifelse(
    sce$orig.ident %in% oe_samples,
    paste0("OE_rep", sce$pseudo_rep),
    paste0(sce$orig.ident, "_rep", sce$pseudo_rep))

  #----------------------
  # 5. Aggregate to pseudobulk
  #----------------------
  pb <- aggregateAcrossCells(
    sce,
    ids = DataFrame(
      pb_id      = sce$pb_id,
      orig.ident = sce$orig.ident,
      condition  = sce$condition,
      pseudo_rep = sce$pseudo_rep
    )
  )
  colnames(pb) <- pb$pb_id

  # Save raw counts
  assay(pb, "counts_raw") <- assay(pb, "counts")
  pb$libsize <- colSums(assay(pb, "counts_raw"))
  
  #----------------------
  # 6. Normalize 
  #----------------------
  pb <- logNormCounts(pb)

  
  #----------------------
  # 8. Return everything
  #----------------------
  return(pb)
}

```



diagnostic
```{r}
pb_sce <- pseudobulk_celltype(seu, celltype="TrEC")
#pb_sce <- runPCA(pb_sce, exprs_values = "logcounts", ncomponents = 5)
```


```{r}
as.data.frame(colData(pb_sce)) %>%
  arrange(ncells) %>%
  select(pb_id, orig.ident, condition, pseudo_rep, ncells)

```




```{r}
plotReducedDim(pb_sce, "PCA", colour_by = "condition", point_size=4) +theme(
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 15)
    )
plotReducedDim(pb_sce, "PCA", colour_by = "orig.ident", point_size=4) +theme(
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 15)
    )
plotReducedDim(pb_sce, "PCA", colour_by = "pseudo_rep", point_size=4) +theme(
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 15)
    )
plotReducedDim(pb_sce, "PCA", colour_by = "ncells", point_size=4) +theme(
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 15)
    )
plotReducedDim(pb_sce, "PCA", colour_by = "libsize", point_size=4) +theme(
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 15)
    )
#plotReducedDim(pb_sce, "PCA", colour_by = "cell_annot")
```

```{r, fig.width=13, fig.height=6}
plotReducedDim(pb_sce, "PCA", colour_by = "condition", point_size=6) +theme(
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 15)
    ) +plotReducedDim(pb_sce, "PCA", colour_by = "orig.ident", point_size=6) +theme(
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 15)
    )
```

```{r}
mat <- assay(pb_sce, "logcounts")
batch <- pb_sce$orig.ident

mat_corrected <- removeBatchEffect(mat, batch=batch)
pc <- prcomp(t(mat_corrected))
df <- data.frame(
  PC1 = pc$x[,1],
  PC2 = pc$x[,2],
  condition = pb_sce$condition,
  orig.ident = pb_sce$orig.ident
)

#library(ggplot2)
ggplot(df, aes(PC1, PC2, color = condition, shape = orig.ident)) +
  geom_point(size=4) + 
  theme(legend.text = element_text(size = 14),
        legend.title = element_text(size = 15))

```

```{r}
seu <- subset(seu, subset = orig.ident != "COUPOE-PP-POS1")
table(seu$cluster_annot, seu$condition)


```


```{r}

#from pb make metadata,design matrix, dge list , filter, norm,disp and GLM
dge_prepare <- function(pb) {
  # library(edgeR)
  # library(dplyr)
  
  # Metadata extraction
  meta <- as.data.frame(colData(pb)) %>%
    dplyr::select(pb_id, condition, pseudo_rep)
  
  meta$condition <- factor(meta$condition)
  meta$pseudo_rep <- factor(meta$pseudo_rep)
  
  # Design matrix (no intercept)
  design <- model.matrix(~ 0 + condition , data = meta)
  
  # Build DGEList using raw pseudobulk counts
  y <- DGEList(
    counts = assay(pb, "counts_raw"),
    samples = meta
  )
  
  # Filter lowly expressed genes
  keep <- filterByExpr(y, design = design)
  y <- y[keep, , keep.lib.sizes = FALSE]
  
  # Normalize (TMM)
  y <- calcNormFactors(y)
  
  # Estimate dispersions
  y <- estimateDisp(y, design)
  
  # Fit GLM
  fit <- glmFit(y, design)
  
  # Return everything needed for contrasts
  return(list(
    fit = fit,
    design = design,
    y = y,
    meta = meta
  ))
}

```



##DGE
```{r}
pb_pcv <- pseudobulk_celltype(seu, celltype = "CapEC1")
pcv_dge <- dge_prepare(pb_pcv)


```


#pos fc means genes are higher in oe/ko/front neg means higher in wt 
```{r Contrast}
fit <- pcv_dge$fit
design <- pcv_dge$design

# 1. OE vs WT
c_oe_wt <- makeContrasts(conditionOE - conditionWT, levels = design)
res_oe_wt <- glmLRT(fit, contrast=c_oe_wt)
deg_oe_wt <- topTags(res_oe_wt, n = Inf)$table

# 2. KO vs WT
c_ko_wt <- makeContrasts(conditionKO - conditionWT, levels = design)
res_ko_wt <- glmLRT(fit,contrast= c_ko_wt)
deg_ko_wt <- topTags(res_ko_wt, n = Inf)$table

# 3. CRE_CTRL vs WT
c_cc_wt <- makeContrasts(conditionCRE_CTRL - conditionWT, levels = design)
res_cc_wt <- glmLRT(fit,contrast= c_cc_wt)
deg_cc_wt <- topTags(res_cc_wt, n = Inf)$table

# 4. OE vs KO
c_oe_ko <- makeContrasts(conditionOE - conditionKO, levels = design)
res_oe_ko <- glmLRT(fit,contrast= c_oe_ko)
deg_oe_ko <- topTags(res_oe_ko, n = Inf)$table
# 
# # 5. OE vs CRE_CTRL
# c_oe_cc <- makeContrasts(conditionOE - conditionCRE_CTRL, levels = design)
# res_oe_cc <- glmLRT(fit, contrast=c_oe_cc)
# deg_oe_cc <- topTags(res_oe_cc, n = Inf)$table
# 
# # 6. KO vs CRE_CTRL
# c_ko_cc <- makeContrasts(conditionKO - conditionCRE_CTRL, levels = design)
# res_ko_cc <- glmLRT(fit,contrast= c_ko_cc)
# deg_ko_cc <- topTags(res_ko_cc, n = Inf)$table


```


OE WT



OE WT

log2fc = higher in oe, neg in wt, cpm is higher for more abundant gene, lr =test stat, bigger stronger evidence

```{r}
head(deg_oe_wt)
```


```{r}
# 
write.csv(deg_oe_wt,
          file.path(celltype_dir, "deg_oe_wt.csv"),
          row.names = TRUE)

# How many significant genes?
sig_genes <- deg_oe_wt %>% filter(FDR < 0.05)
nrow(sig_genes)
# Output: xxx significant genes

# How many up vs down?
table(sig_genes$logFC > 0)
# FALSE (down in OE): xxx
# TRUE (up in OE): xxx
```
[1] 185

FALSE  TRUE 
   77   108 
```{r}
#ibrary(dplyr)

# Top 20 OE-upregulated genes
top_up <- deg_oe_wt %>%
  filter(FDR < 0.05, logFC > 1.2) %>%
  arrange(desc(logFC)) %>% 
  head(20)

# Top 20 WT-upregulated genes (down in OE)
top_down <- deg_oe_wt %>%
  filter(FDR < 0.05, logFC < 1.2) %>%
  arrange(logFC) %>% 
  head(20)
write.csv(top_up,
          file.path(celltype_dir, "deg_oe_wt_top20_up.csv"),
          row.names = TRUE)

write.csv(top_down,
          file.path(celltype_dir, "deg_oe_wt_top20_down.csv"),
          row.names = TRUE)

```




Viz

```{r,fig.width=9,fig.height=7}
p <- EnhancedVolcano(
  deg_oe_wt,
  lab = rownames(deg_oe_wt),
  x = 'logFC',
  y = 'FDR',
  pCutoff = 0.05,
  FCcutoff = 1.2,     # log2 fold change cutoff of ±1
  pointSize = 2.0,
  labSize = 5,
  #col = c('grey30', 'grey30', 'royalblue', 'red2'),
  title = 'OE vs WT (CapEC1)',
  subtitle = 'DEG Volcano Plot',
  caption = 'Right = Upregulated | Left = Downregulated',
  legendPosition = 'right'
)
p
save_pub(plot=p,filename= "deg_oe_wt_EVplot", width = 9,height=7)
```


```{r, fig.width=9, fig.height=7}
# Genes you care about
goi <- c("Madcam1", "St6gal1",     "Nr2f2" ,"Nkx2-3", "Hey1",  "Cadm1","Clec2d",   "Ccl21","Sele", "Ephb4" ,"Selp", "Vcam1", "Glycam1", "Icam1")
         #addressin                 #tf                                  #general pcv/hev genes

p <- EnhancedVolcano(
  deg_oe_wt,
  lab = rownames(deg_oe_wt),
  x = 'logFC',
  y = 'FDR',
  pCutoff = 0.05,
  FCcutoff = 1.2,
  selectLab = goi,              # Only label these genes
  drawConnectors = TRUE,         # Draw lines to labels
  widthConnectors = 0.5,
  colConnectors = 'black',
  title = 'OE vs WT - Adhesion Molecules CapEC1'
)

save_pub(plot=p,filename= "deg_oe_wt_goi_EVplot", width = 9,height=7)
```



```{r}
# Subset by cell type (example: CRP)
crp <- subset(seu, idents = "CRP")

# Violin plot of a feature split by condition
VlnPlot(
  object = crp,
  features = "Cd31",
  group.by = "condition",
  pt.size = 0
)

```


```{r, fig.width=9, fig.height=7}
# Genes you care about
goi <- c("Madcam1", "St6gal1" ,"Nkx2-3", "Ccl21a","Ubd", "Fut7","Ltb", "Oit1", "Clu" ,"Vwf" ,"Selp", "Sele", "Vcam1", "Glycam1", "Icam1","Nr2f2")
         #addressin                                                   

p <- EnhancedVolcano(
  deg_oe_wt,
  lab = rownames(deg_oe_wt),
  x = 'logFC',
  y = 'FDR',
  pCutoff = 0.05,
  FCcutoff = 1.2,
  selectLab = goi,              # Only label these genes
  drawConnectors = TRUE,         # Draw lines to labels
  widthConnectors = 0.5,
  colConnectors = 'black',
  title = 'Is There Reprogramming of CapEC1 to HEV? OE vs WT'
)

save_pub(plot=p,filename= "degtrecreprog", width = 9,height=7)
```




```{r,fig.height=9,fig.width=11}
library(pheatmap)

# Get top 50 DEGs
top_degs <- deg_oe_wt %>%
  filter(FDR < 0.05) %>%
  arrange(FDR) %>%
  head(50)

# Extract normalized counts for these genes
library(edgeR)
norm_counts <- cpm(pcv_dge$y, log = TRUE)  # Log-CPM normalized
top_genes_mat <- norm_counts[rownames(top_degs), ]

# Create heatmap
p <- pheatmap(
  top_genes_mat,
  scale = "row",              # Z-score scaling
  cluster_rows = TRUE,
  cluster_cols = TRUE,
  show_rownames = TRUE,
  show_colnames = TRUE,
  annotation_col = data.frame(
    Condition = pcv_dge$meta$condition,
    row.names = colnames(top_genes_mat)
  ),
  main = "Top 50 DEGs for CapEC1: OE vs WT"
)

p
save_pub(plot=p, filename ="heatmap_deg_oe_wt",width = 11, height=9)
```


GO enrichment using ORA
take significant deg, convert to entrez id and check if the sig genes are over represented in GO (biological processes)

```{r}
# Significant DE genes
sig_deg <- deg_oe_wt %>% 
  filter(FDR < 0.05) 

# Up and down separately (optional)
up_deg <- sig_deg %>% filter(logFC > 0)
down_deg <- sig_deg %>% filter(logFC < 0)

# 
# library(AnnotationDbi)
# 
# library(org.Mm.eg.db)
# library(clusterProfiler)

gene_symbols <- rownames(sig_deg)

entrez_ids <- bitr(
  gene_symbols,
  fromType = "SYMBOL",
  toType = "ENTREZID",
  OrgDb = org.Mm.eg.db
)

sig_deg_entrez <- sig_deg %>% 
  mutate(SYMBOL = rownames(sig_deg)) %>%
  inner_join(entrez_ids, by = "SYMBOL")


ego_bp <- enrichGO(
  gene          = sig_deg_entrez$ENTREZID,
  OrgDb         = org.Mm.eg.db,
  keyType       = "ENTREZID",
  ont           = "BP",
  pAdjustMethod = "fdr",
  pvalueCutoff  = 0.05,
  readable      = TRUE
)


write.csv(ego_bp,
          file.path(celltype_dir, "ego_bp_deg_oe_wt.csv"),
          row.names = FALSE)


head(ego_bp)

```

```{r, fig.height=9, fig.width=14}
#this is by padj
b <- barplot(ego_bp, showCategory = 20)
d <- dotplot(ego_bp, showCategory = 20)
save_pub(plot =b, filename="barplot_GO_deg_oe_wt", width=14, height=9)
save_pub(plot =d, filename="dotplot_GO_deg_oe_wt", width=14, height=9)
```

```{r, fig.height=14}
d <- dotplot(ego_bp, showCategory = 10) +
  theme(
    axis.text.y  = element_text(size = 24, face = "bold"),  # bigger + bold
    axis.text.x  = element_text(size = 22, face = "bold"),  # bigger + bold
    legend.title = element_text(size = 22, face = "bold"),  # legend title
    legend.text  = element_text(size = 22),axis.text.y.left = element_text(margin = margin(r = 15))                 # legend labels
  )

d
save_pub(plot=d, filename ="oewtcapecora",width = 12, height=14)
```



fsgea
1.Ranks all genes by logfc more up in oe at top and most up in wt at the bottom
2. MsigBS gene set c3 is TF target geens, C2 is kegg, reactome etc
3.Filter for ets related pathways
4.NES (Normalized Enrichment Score):

Positive NES: Gene set enriched in upregulated genes (OE has more)
Negative NES: Gene set enriched in downregulated genes (WT has more)
Magnitude matters: |NES| > 2 is strong, > 1.5 is moderate


```{r}
ranks <- deg_oe_wt$logFC
names(ranks) <- rownames(deg_oe_wt)
ranks <- sort(ranks, decreasing = TRUE)

#msig_c3 <- msigdbr(species = "Mus musculus", collection = "C3")

c3_tft <- split(msig_c3$gene_symbol, msig_c3$gs_name)

fgsea_ets <- fgsea(
  pathways = c3_tft,
  stats = ranks,
  minSize = 10,
  maxSize = 500
)

fgsea_ets$leadingEdge <- sapply(fgsea_ets$leadingEdge, paste, collapse = ",")

write.csv(
  fgsea_ets,
  file.path(celltype_dir, "fgsea_ets_all_before_grep_deg_oe_wt.csv"),
  row.names = FALSE
)

# ------------------------------------------------
fgsea_ets <- fgsea_ets[
  grep("^(ETS|ERG|PEA3|ETV2|TCF|ERF|PDEF|ELF|ESE|TEL|SPI|ELG)", fgsea_ets$pathway),
]


write.csv(
  fgsea_ets,
  file.path(celltype_dir, "fgsea_ets_only_deg_oe_wt.csv"),
  row.names = FALSE
)

# FDR filter + order (your minimal fixes)
#fgsea_ets <- fgsea_ets[fgsea_ets$padj < 0.26, ]
fgsea_ets <- fgsea_ets[order(fgsea_ets$padj), ]

top_ets <- head(fgsea_ets, 20)
top_ets



```
```{r}
p <- ggplot(top_ets, aes(x = NES, y = reorder(pathway, NES))) +
  geom_point(aes(size = -log10(padj), color = NES)) +
  scale_color_gradient2(low = "blue", high = "red", mid = "white") +
  labs(
    title = "ETS-related enrichment CapEC1 (OE vs WT)",
    x = "Normalized Enrichment Score (NES)",
    y = "ETS target sets",
    size = "-log10(FDR)"
  ) +
  theme_bw()

save_pub(plot=p, filename="ets_top20_dotplot_deg_oe_wt", width=8, height=8)
```


Absolutely no relevant enrichment in Capec1 oe v wt!



```{r}
ranks <- deg_oe_wt$logFC
names(ranks) <- rownames(deg_oe_wt)
ranks <- sort(ranks, decreasing = TRUE)

fgsea_kegg_full <- fgsea(
  pathways = kegg_gs,
  stats    = ranks,
  minSize  = 10,
  maxSize  = 500
)
fgsea_kegg_full <- fgsea_kegg_full[order(fgsea_kegg_full$padj), ]

## 5. Count how many ETS TFs each KEGG pathway contains
fgsea_kegg_full$n_ets <- sapply(
  fgsea_kegg_full$pathway,
  function(p) length(intersect(kegg_gs[[p]], ets_family_genes))
)


## 6. Keep KEGG pathways that:
##    (a) are enriched (FDR < 0.25)
##    (b) contain >= 1 ETS TF gene
ets_kegg <- subset(fgsea_kegg_full, padj < 0.25 & n_ets > 0)
ets_kegg <- ets_kegg[order(ets_kegg$padj), ]

ets_kegg

ets_kegg$leadingEdge <- sapply(ets_kegg$leadingEdge, paste, collapse = ",")
write.csv(
  ets_kegg,
  file.path(celltype_dir, "ets_kegg_fsgea_deg_oe_wt.csv"),
  row.names = FALSE
)
```
```{r}


# convert SYMBOL -> ENTrez
mapped_ids <- mapIds(
  org.Mm.eg.db,
  keys = names(ranks),
  column = "ENTREZID",
  keytype = "SYMBOL",
  multiVals = "first"
)

ranks_entrez <- ranks
names(ranks_entrez) <- mapped_ids
ranks_entrez <- ranks_entrez[!is.na(names(ranks_entrez))]

# old <- getwd()
# setwd(celltype_dir)
# 
# pathview(
#   gene.data  = ranks_entrez,# will create this below
#   pathway.id = "mmu04218",     # your enriched KEGG pathway
#   species    = "mmu",
#   out.suffix =  "OE_vs_WT"
# )
# setwd(old)



```

nothing

```{r}

old <- getwd()
setwd(celltype_dir)
kegg_ids <- ets_kegg$pathway   # the enriched KEGG IDs you filtered

for (pid in kegg_ids) {
  message("Running Pathview for: ", pid, " (OE_vs_WT)")

  pathview(
    gene.data  = ranks_entrez,
    pathway.id = pid,
    species    = "mmu",
    out.suffix = paste0("OE_vs_WT_hev", pid)
  )
}

setwd(old)
```

What it does: Creates dot plot of top 20 ETS-related pathways
How to interpret:

X-axis: NES (right = upregulated in OE, left = upregulated in WT)
Dot size: Significance (bigger = more significant)
Dot color: Same as NES (red = up in OE, blue = up in WT)



KO WT

```{r}
# 
write.csv(deg_ko_wt,
           file.path(celltype_dir, "deg_ko_wt.csv"),
           row.names = TRUE)

# How many significant genes?
sig_genes <- deg_ko_wt %>% filter(FDR < 0.05)
nrow(sig_genes)
# Output: xxx significant genes

# How many up vs down?
table(sig_genes$logFC > 0)
# FALSE (down in OE): xxx
# TRUE (up in OE): xxx
```
[1] 314

FALSE  TRUE 
  159   155 
  
  
```{r}
#ibrary(dplyr)

# Top 20 ko-upregulated genes
top_up <- deg_ko_wt %>%
  filter(FDR < 0.05, logFC > 1.2) %>%
  arrange(desc(logFC)) %>% 
  head(20)

# Top 20 WT-upregulated genes (down in ko)
top_down <- deg_ko_wt %>%
  filter(FDR < 0.05, logFC < 1.2) %>%
  arrange(logFC) %>% 
  head(20)
write.csv(top_up,
          file.path(celltype_dir, "deg_ko_wt_top20_up.csv"),
          row.names = TRUE)

write.csv(top_down,
          file.path(celltype_dir, "deg_ko_wt_top20_down.csv"),
          row.names = TRUE)

```

viz

```{r,fig.width=9,fig.height=7}
p <- EnhancedVolcano(
  deg_ko_wt,
  lab = rownames(deg_ko_wt),
  x = 'logFC',
  y = 'FDR',
  pCutoff = 0.05,
  FCcutoff = 1.2,     # log2 fold change cutoff of ±1
  pointSize = 2.0,
  labSize = 5,
  #col = c('grey30', 'grey30', 'royalblue', 'red2'),
  title = 'KO vs WT (CapEC1)',
  subtitle = 'DEG Volcano Plot',
  caption = 'Right = Upregulated | Left = Downregulated',
  legendPosition = 'right'
)
p
save_pub(plot=p,filename= "deg_ko_wt_EVplot", width = 9,height=7)
```

```{r, fig.width=9, fig.height=7}
# Genes you care about
goi <- c("Madcam1", "St6gal1","Ly6c",     "Nr2f2" ,"Nkx2-3", "Hey1",     "Ccl21","Sele", "Ephb4" ,"Selp", "Vcam1", "Glycam1", "Icam1", "Ltb")
         #addressin                 #tf                                  #general pcv/hev genes

p <- EnhancedVolcano(
  deg_ko_wt,
  lab = rownames(deg_ko_wt),
  x = 'logFC',
  y = 'FDR',
  pCutoff = 0.05,
  FCcutoff = 1.2,
  selectLab = goi,              # Only label these genes
  drawConnectors = TRUE,         # Draw lines to labels
  widthConnectors = 0.5,
  colConnectors = 'black',
  title = 'KO vs WT - Adhesion Molecules CapEC1'
)

save_pub(plot=p,filename= "deg_ko_wt_goi_EVplot", width = 9,height=7)
```

The KO doesnt even seem to have been knoacked out properly!! Or just that we are not observing enough of each in it ot be significant. So we will NOT proceed with the rest of the analysis.

```{r,fig.height=9,fig.width=11}
library(pheatmap)

# Get top 50 DEGs
top_degs <- deg_ko_wt %>%
  filter(FDR < 0.05) %>%
  arrange(FDR) %>%
  head(50)

# Extract normalized counts for these genes
#library(edgeR)
norm_counts <- cpm(pcv_dge$y, log = TRUE)  # Log-CPM normalized
top_genes_mat <- norm_counts[rownames(top_degs), ]

# Create heatmap
p <- pheatmap(
  top_genes_mat,
  scale = "row",              # Z-score scaling
  cluster_rows = TRUE,
  cluster_cols = TRUE,
  show_rownames = TRUE,
  show_colnames = TRUE,
  annotation_col = data.frame(
    Condition = pcv_dge$meta$condition,
    row.names = colnames(top_genes_mat)
  ),
  main = "Top 50 DEGs for CapEC1: KO vs WT"
)

p
save_pub(plot=p, filename ="heatmap_deg_ko_wt",width = 11, height=9)
```


GO for KO vs WT


```{r}
# Significant DE genes
sig_deg <- deg_ko_wt %>% 
  filter(FDR < 0.05) 

# Up and down separately (optional)
up_deg <- sig_deg %>% filter(logFC > 0)
down_deg <- sig_deg %>% filter(logFC < 0)

# 
# library(AnnotationDbi)
# 
# library(org.Mm.eg.db)
# library(clusterProfiler)

gene_symbols <- rownames(sig_deg)

entrez_ids <- bitr(
  gene_symbols,
  fromType = "SYMBOL",
  toType = "ENTREZID",
  OrgDb = org.Mm.eg.db
)

sig_deg_entrez <- sig_deg %>% 
  mutate(SYMBOL = rownames(sig_deg)) %>%
  inner_join(entrez_ids, by = "SYMBOL")


ego_bp <- enrichGO(
  gene          = sig_deg_entrez$ENTREZID,
  OrgDb         = org.Mm.eg.db,
  keyType       = "ENTREZID",
  ont           = "BP",
  pAdjustMethod = "fdr",
  pvalueCutoff  = 0.05,
  readable      = TRUE
)

head(ego_bp)

write.csv(ego_bp,
          file.path(celltype_dir, "ego_bp_deg_ko_wt.csv"),
          row.names = FALSE)
```

```{r, fig.height=9, fig.width=14}
#this is by padj
b <- barplot(ego_bp, showCategory = 20)
d <- dotplot(ego_bp, showCategory = 20)
save_pub(plot =b, filename="barplot_GO_deg_ko_wt", width=14, height=9)
save_pub(plot =d, filename="dotplot_GO_deg_ko_wt", width=14, height=9)
```

```{r}
d <- dotplot(ego_bp, showCategory = 10) +
  theme(
    axis.text.y  = element_text(size = 24, face = "bold"),  # bigger + bold
    axis.text.x  = element_text(size = 22, face = "bold"),  # bigger + bold
    legend.title = element_text(size = 22, face = "bold"),  # legend title
    legend.text  = element_text(size = 22),axis.text.y.left = element_text(margin = margin(r = 15))                 # legend labels
  )

d
save_pub(plot=d, filename ="kowtcapecora",width = 12, height=14)
```



fsgea

```{r}
ranks <- deg_ko_wt$logFC
names(ranks) <- rownames(deg_ko_wt)
ranks <- sort(ranks, decreasing = TRUE)

#msig_c3 <- msigdbr(species = "Mus musculus", collection = "C3")

c3_tft <- split(msig_c3$gene_symbol, msig_c3$gs_name)

fgsea_ets <- fgsea(
  pathways = c3_tft,
  stats = ranks,
  minSize = 10,
  maxSize = 500
)

fgsea_ets$leadingEdge <- sapply(fgsea_ets$leadingEdge, paste, collapse = ",")

write.csv(
  fgsea_ets,
  file.path(celltype_dir, "fgsea_ets_all_before_grep_deg_ko_wt.csv"),
  row.names = FALSE
)

# ------------------------------------------------
fgsea_ets <- fgsea_ets[
  grep("^(ETS|ERG|PEA3|ETV2|TCF|ERF|PDEF|ELF|ESE|TEL|SPI|ELG)", fgsea_ets$pathway),
]


# FDR filter + order (your minimal fixes)
#fgsea_ets <- fgsea_ets[fgsea_ets$padj < 0.26, ]
fgsea_ets <- fgsea_ets[order(fgsea_ets$padj), ]


write.csv(
  fgsea_ets,
  file.path(celltype_dir, "fgsea_ets_only_deg_ko_wt.csv"),
  row.names = FALSE
)


top_ets <- head(fgsea_ets, 20)
top_ets
p <- ggplot(top_ets, aes(x = NES, y = reorder(pathway, NES))) +
  geom_point(aes(size = -log10(padj), color = NES)) +
  scale_color_gradient2(low = "blue", high = "red", mid = "white") +
  labs(
    title = "ETS-related enrichment CapEC1 (KO vs WT)",
    x = "Normalized Enrichment Score (NES)",
    y = "ETS target sets",
    size = "-log10(FDR)"
  ) +
  theme_bw()

save_pub(plot=p, filename="ets_top20_dotplot_deg_ko_wt", width=8, height=8)


```
```{r}
top_ets <- head(fgsea_ets, 20)
top_ets
p <- ggplot(top_ets, aes(x = NES, y = reorder(pathway, NES))) +
  geom_point(aes(size = -log10(padj), color = NES)) +
  scale_color_gradient2(low = "blue", high = "red", mid = "white") +
  labs(
    title = "ETS-related enrichment CapEC1 (KO vs WT)",
    x = "Normalized Enrichment Score (NES)",
    y = "ETS target sets",
    size = "-log10(FDR)"
  ) +
  theme_bw()

save_pub(plot=p, filename="ets_top20_dotplot_deg_ko_wt", width=8, height=8)

```


```{r}
fgsea_kegg_full <- fgsea(
  pathways = kegg_gs,
  stats    = ranks,
  minSize  = 10,
  maxSize  = 500
)
fgsea_kegg_full <- fgsea_kegg_full[order(fgsea_kegg_full$padj), ]

## 5. Count how many ETS TFs each KEGG pathway contains
fgsea_kegg_full$n_ets <- sapply(
  fgsea_kegg_full$pathway,
  function(p) length(intersect(kegg_gs[[p]], ets_family_genes))
)


## 6. Keep KEGG pathways that:
##    (a) are enriched (FDR < 0.25)
##    (b) contain >= 1 ETS TF gene
ets_kegg <- subset(fgsea_kegg_full, padj < 0.25 & n_ets > 0)
ets_kegg <- ets_kegg[order(ets_kegg$padj), ]

ets_kegg

ets_kegg$leadingEdge <- sapply(ets_kegg$leadingEdge, paste, collapse = ",")
write.csv(
  ets_kegg,
  file.path(celltype_dir, "ets_kegg_fsgea_deg_ko_wt.csv"),
  row.names = FALSE
)
```


```{r}

old <- getwd()
setwd(celltype_dir)
kegg_ids <- ets_kegg$pathway   # the enriched KEGG IDs you filtered

for (pid in kegg_ids) {
  message("Running Pathview for: ", pid, " (KO_vs_WT)")

  pathview(
    gene.data  = ranks_entrez,
    pathway.id = pid,
    species    = "mmu",
    out.suffix = paste0("KO_vs_WT_capec1", pid)
  )
}

setwd(old)
```



What it does: Creates dot plot of top 20 ETS-related pathways
How to interpret:

X-axis: NES (right = upregulated in OE, left = upregulated in WT)
Dot size: Significance (bigger = more significant)
Dot color: Same as NES (red = up in OE, blue = up in WT)



OE VS KO


```{r}

write.csv(deg_oe_ko,
          file.path(celltype_dir, "deg_oe_ko.csv"),
          row.names = TRUE)

# How many significant genes?
sig_genes <- deg_oe_ko %>% filter(FDR < 0.05)
nrow(sig_genes)
# Output: xxx significant genes

# How many up vs down?
table(sig_genes$logFC > 0)
# FALSE (down in OE): xxx
# TRUE (up in OE): xxx
```
1] 283

FALSE  TRUE 
  114   169 

```{r}
#ibrary(dplyr)

# Top 20 OE-upregulated genes
top_up <- deg_oe_ko %>%
  filter(FDR < 0.05, logFC > 1.2) %>%
  arrange(desc(logFC)) %>% 
  head(20)

# Top 20 WT-upregulated genes (down in OE)
top_down <- deg_oe_ko %>%
  filter(FDR < 0.05, logFC < 1.2) %>%
  arrange(logFC) %>% 
  head(20)
write.csv(top_up,
          file.path(celltype_dir, "deg_oe_ko_top20_up.csv"),
          row.names = TRUE)

write.csv(top_down,
          file.path(celltype_dir, "deg_oe_ko_top20_down.csv"),
          row.names = TRUE)

```


```{r}
# library(DT)
# 
# # Interactive searchable table
# datatable(deg_oe_wt, 
#           filter = 'top',
#           options = list(pageLength = 50))
```


Viz

```{r,fig.width=9,fig.height=7}
p <- EnhancedVolcano(
  deg_oe_ko,
  lab = rownames(deg_oe_ko),
  x = 'logFC',
  y = 'FDR',
  pCutoff = 0.05,
  FCcutoff = 1.2,     # log2 fold change cutoff of ±1
  pointSize = 2.0,
  labSize = 5,
  #col = c('grey30', 'grey30', 'royalblue', 'red2'),
  title = 'OE vs KO (CapEC1)',
  subtitle = 'DEG Volcano Plot',
  caption = 'Right = Upregulated | Left = Downregulated',
  legendPosition = 'right'
)
p
save_pub(plot=p,filename= "deg_oe_ko_EVplot", width = 9,height=7)
```


```{r, fig.width=9, fig.height=7}
# Genes you care about
goi <- c("Madcam1", "St6gal1",     "Nr2f2" ,"Nkx2-3", "Hey1",     "Ccl21","Sele", "Ephb4" ,"Selp", "Vcam1", "Glycam1", "Icam1")
         #addressin                 #tf                                  #general pcv/hev genes

p <- EnhancedVolcano(
  deg_oe_ko,
  lab = rownames(deg_oe_ko),
  x = 'logFC',
  y = 'FDR',
  pCutoff = 0.05,
  FCcutoff = 1.2,
  selectLab = goi,              # Only label these genes
  drawConnectors = TRUE,         # Draw lines to labels
  widthConnectors = 0.5,
  colConnectors = 'black',
  title = 'OE vs KO - Adhesion Molecules'
)

save_pub(plot=p,filename= "deg_oe_ko_goi_EVplot", width = 9,height=7)
```


```{r,fig.height=9,fig.width=11}
library(pheatmap)

# Get top 50 DEGs
top_degs <- deg_oe_ko %>%
  filter(FDR < 0.05) %>%
  arrange(FDR) %>%
  head(50)

# Extract normalized counts for these genes
library(edgeR)
norm_counts <- cpm(pcv_dge$y, log = TRUE)  # Log-CPM normalized
top_genes_mat <- norm_counts[rownames(top_degs), ]

# Create heatmap
p <- pheatmap(
  top_genes_mat,
  scale = "row",              # Z-score scaling
  cluster_rows = TRUE,
  cluster_cols = TRUE,
  show_rownames = TRUE,
  show_colnames = TRUE,
  annotation_col = data.frame(
    Condition = pcv_dge$meta$condition,
    row.names = colnames(top_genes_mat)
  ),
  main = "Top 50 DEGs for CAPEC1: OE vs KO"
)

p
save_pub(plot=p, filename ="heatmap_deg_oe_ko",width = 11, height=9)
```


GO enrichment using ORA
take significant deg, convert to entrez id and check if the sig genes are over represented in GO (biological processes)

```{r}
# Significant DE genes
sig_deg <- deg_oe_ko %>% 
  filter(FDR < 0.05) 

# Up and down separately (optional)
up_deg <- sig_deg %>% filter(logFC > 0)
down_deg <- sig_deg %>% filter(logFC < 0)

# 
# library(AnnotationDbi)
# 
# library(org.Mm.eg.db)
# library(clusterProfiler)

gene_symbols <- rownames(sig_deg)

entrez_ids <- bitr(
  gene_symbols,
  fromType = "SYMBOL",
  toType = "ENTREZID",
  OrgDb = org.Mm.eg.db
)

sig_deg_entrez <- sig_deg %>% 
  mutate(SYMBOL = rownames(sig_deg)) %>%
  inner_join(entrez_ids, by = "SYMBOL")


ego_bp <- enrichGO(
  gene          = sig_deg_entrez$ENTREZID,
  OrgDb         = org.Mm.eg.db,
  keyType       = "ENTREZID",
  ont           = "BP",
  pAdjustMethod = "fdr",
  pvalueCutoff  = 0.05,
  readable      = TRUE
)


write.csv(ego_bp,
          file.path(celltype_dir, "ego_bp_deg_oe_ko.csv"),
          row.names = FALSE)
head(ego_bp)

```

```{r, fig.height=9, fig.width=14}
#this is by padj
b <- barplot(ego_bp, showCategory = 20)
d <- dotplot(ego_bp, showCategory = 20)
save_pub(plot =b, filename="barplot_GO_deg_oe_ko", width=14, height=9)
save_pub(plot =d, filename="dotplot_GO_deg_oe_ko", width=14, height=9)
```



fsgea
1.Ranks all genes by logfc more up in oe at top and most up in wt at the bottom
2. MsigBS gene set c3 is TF target geens, C2 is kegg, reactome etc
3.Filter for ets related pathways
4.NES (Normalized Enrichment Score):

Positive NES: Gene set enriched in upregulated genes (OE has more)
Negative NES: Gene set enriched in downregulated genes (WT has more)
Magnitude matters: |NES| > 2 is strong, > 1.5 is moderate


```{r}
ranks <- deg_oe_ko$logFC
names(ranks) <- rownames(deg_oe_ko)
ranks <- sort(ranks, decreasing = TRUE)

msig_c3 <- msigdbr(species = "Mus musculus", collection = "C3")

c3_tft <- split(msig_c3$gene_symbol, msig_c3$gs_name)

fgsea_ets <- fgsea(
  pathways = c3_tft,
  stats = ranks,
  minSize = 10,
  maxSize = 500
)

fgsea_ets$leadingEdge <- sapply(fgsea_ets$leadingEdge, paste, collapse = ",")

write.csv(
  fgsea_ets,
  file.path(celltype_dir, "fgsea_ets_all_before_grep_deg_oe_ko.csv"),
  row.names = FALSE
)

# ------------------------------------------------
fgsea_ets <- fgsea_ets[
  grep("^(ETS|ERG|PEA3|ETV2|TCF|ERF|PDEF|ELF|ESE|TEL|SPI|ELG)", fgsea_ets$pathway),
]

write.csv(
  fgsea_ets,
  file.path(celltype_dir, "fgsea_ets_only_deg_oe_ko.csv"),
  row.names = FALSE
)

# FDR filter + order (your minimal fixes)
#fgsea_ets <- fgsea_ets[fgsea_ets$padj < 0.26, ]
fgsea_ets <- fgsea_ets[order(fgsea_ets$padj), ]

top_ets <- head(fgsea_ets, 20)
top_ets
p <- ggplot(top_ets, aes(x = NES, y = reorder(pathway, NES))) +
  geom_point(aes(size = -log10(padj), color = NES)) +
  scale_color_gradient2(low = "blue", high = "red", mid = "white") +
  labs(
    title = "ETS-related enrichment CapEC2 (OE vs KO)",
    x = "Normalized Enrichment Score (NES)",
    y = "ETS target sets",
    size = "-log10(FDR)"
  ) +
  theme_bw()

save_pub(plot=p, filename="ets_top20_dotplot_deg_oe_ko", width=8, height=8)


```
TCF4_Q5 Enriched in OE.

```{r}
fgsea_kegg_full <- fgsea(
  pathways = kegg_gs,
  stats    = ranks,
  minSize  = 10,
  maxSize  = 500
)
fgsea_kegg_full <- fgsea_kegg_full[order(fgsea_kegg_full$padj), ]

## 5. Count how many ETS TFs each KEGG pathway contains
fgsea_kegg_full$n_ets <- sapply(
  fgsea_kegg_full$pathway,
  function(p) length(intersect(kegg_gs[[p]], ets_family_genes))
)


## 6. Keep KEGG pathways that:
##    (a) are enriched (FDR < 0.25)
##    (b) contain >= 1 ETS TF gene
ets_kegg <- subset(fgsea_kegg_full, padj < 0.25 & n_ets > 0)
ets_kegg <- ets_kegg[order(ets_kegg$padj), ]

ets_kegg

ets_kegg$leadingEdge <- sapply(ets_kegg$leadingEdge, paste, collapse = ",")
write.csv(
  ets_kegg,
  file.path(celltype_dir, "ets_kegg_fsgea_deg_oe_ko.csv"),
  row.names = FALSE)

```
```{r}
# convert SYMBOL -> ENTrez
mapped_ids <- mapIds(
  org.Mm.eg.db,
  keys = names(ranks),
  column = "ENTREZID",
  keytype = "SYMBOL",
  multiVals = "first"
)

ranks_entrez <- ranks
names(ranks_entrez) <- mapped_ids
ranks_entrez <- ranks_entrez[!is.na(names(ranks_entrez))]

```

not doing this
```{r}
kegg_ids <- ets_kegg$pathway   # the enriched KEGG IDs you filtered

for (pid in kegg_ids) {
  message("Running Pathview for: ", pid, " (OE_vs_KO)")

  pathview(
    gene.data  = ranks_entrez,
    pathway.id = pid,
    species    = "mmu",
    out.suffix = paste0("OE_vs_KO_hev", pid)
  )
}

setwd(old)
```

```{r}

old <- getwd()
setwd(celltype_dir)

pathview(
  gene.data  = ranks_entrez,# will create this below
  pathway.id = "mmu04014",     # your enriched KEGG pathway
  species    = "mmu",
  out.suffix =  "OE_vs_KO"
)
setwd(old)
```

What it does: Creates dot plot of top 20 ETS-related pathways
How to interpret:

X-axis: NES (right = upregulated in OE, left = upregulated in WT)
Dot size: Significance (bigger = more significant)
Dot color: Same as NES (red = up in OE, blue = up in WT)


